### 2022-11-21 draft/2020 rejected - JEP General/RandomDeterministicNoise/Title page.tex

\documentclass[12pt]{article}
\usepackage{hyperref}
\usepackage[authoryear, round,sort,comma,numbers]{natbib}
\usepackage{times}
\usepackage{color}
\usepackage{apalike}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage{amsmath}
\usepackage[font={sf,small}]{caption}
\usepackage{amssymb}
\usepackage{float}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\setlength{\textheight}{9.3in}
\setlength{\textwidth}{7in}
\setlength{\footskip}{0.5in}
\setlength{\topmargin}{-0.5in}
\setlength{\headheight}{0.2in}
\setlength{\headsep}{0in}
\setlength{\parindent}{1pc}
\setlength{\oddsidemargin}{-0.25in}
\setlength{\evensidemargin}{-0.25in}
\renewcommand{\baselinestretch}{1.5}

\newcommand*{\email}[1]{%
    \normalsize\href{mailto:#1}{#1}\par
    }

\title{The nature of decision noise in random exploration}

\author[1]{Siyu Wang\thanks{sywangr@email.arizona.edu}}
\author[1,2]{Robert C. Wilson}

\affil[1]{Department of Psychology, University of Arizona, Tucson AZ USA}
\affil[2]{Cognitive Science Program, University of Arizona, Tucson AZ USA}

\date{\today}

\begin{document}
	\maketitle

\end{document}