### 2022-12-21 OLD/oldversions/Wilson_Siyu_RandomDeterministicNoise/codes/GRAPHICALMODEL/EEHorizon_2sigma.tex
add older version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template graphical model %
% Author: Tom Lodewyckx    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% 0. Begin document
%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{pst-all}
\usepackage{com.braju.graphicalmodels}
\usepackage{ifpdf}
\ifpdf
\pdfpaperheight=10.5in
\pdfpaperwidth=5.5in
\else
\special{papersize=10.5in,5.5in}
\fi
\catcode`\@=11%
\pagestyle{empty}
\begin{document}
	%\TeXtoEPSepsi
	\psset{unit=14mm,arrowscale=1.5}
	\SpecialCoor

% 1. Grid
%%%%%%%%%

\begin{pspicture}(4,2)(19,8)
%\showgrid

% 2. Nodes
%%%%%%%%%%

%\rput(1,1){\GM@node[NodeOptions]{name}}
%                 \GM@label[LabelOptions]{name}{label}

% dm and dI
\rput(5,3.25){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{isg}$}
\rput(7,3.25){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{isg}$}

% choices
\rput(6,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{c}}\GM@label[offset=-3mm]{c}{$c_{isgr}$}

% subject parameters
\rput(4,8.5){\GM@node[nodeSize=12mm]{A}}\GM@label[offset=-3mm]{A}{$A_{is}$}
\rput(6,8.5){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$b_{is}$}
\rput(8,8.5){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma^{ran}_{is}$}
\rput(10,8.5){\GM@node[nodeSize=12mm]{epsilon}}\GM@label[offset=-2mm]{epsilon}{$\sigma^{det}_{is}$}

% hyperparameters
\rput(3,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{i}$}
\rput(4.07,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{i}$}

\rput(5.14,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^b_{i}$}
\rput(6.21,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^b_{i}$}

\rput(7.28,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^{ran}_{i}$}
\rput(8.35,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^{ran}_{i}$}

\rput(9.42,10){\GM@node[nodeSize=12mm]{kEpsilon}}\GM@label[offset=-2mm]{kEpsilon}{$k^{det}_{i}$}
\rput(10.49,10){\GM@node[nodeSize=12mm]{lambdaEpsilon}}\GM@label[offset=-2mm]{lambdaEpsilon}{$\lambda^{det}_{i}$}

\rput(8,6){\GM@node[nodeSize=12mm]{sigmar}}\GM@label[offset=-4mm]{sigmar}{$n^{ran}_{isgr}$}
\rput(10,4.5){\GM@node[nodeSize=12mm]{epsilonr}}\GM@label[offset=-3mm]{epsilonr}{$n^{det}_{isg}$}

% differences
%\rput(3,13){\GM@node[nodeSize=12mm]{dmuAOUT}}
%\rput(3,13){\GM@node[nodeSize=10mm]{dmuA}}\GM@label[offset=-3mm]{dmuA}{$\delta\mu^A_{u}$}
%\rput(5,13){\GM@node[nodeSize=12mm]{dkSigmaOUT}}
%\rput(5,13){\GM@node[nodeSize=10mm]{dkSigma}}\GM@label[offset=-3mm]{dkSigma}{$\delta k^\sigma_{u}$}
%\rput(7,13){\GM@node[nodeSize=12mm]{dmubOUT}}
%\rput(7,13){\GM@node[nodeSize=10mm]{dmub}}\GM@label[offset=-3mm]{dmub}{$\delta\mu^b_{u}$}

% 3. Arrows
%%%%%%%%%%%
%
\ncline[arrows=->]{dm}{c}
\ncline[arrows=->]{dI}{c}

\ncline[arrows=->]{A}{c}
\ncline[arrows=->]{sigmar}{c}
\ncline[arrows=->]{b}{c}
\ncline[arrows=->]{epsilonr}{c}

\ncline[arrows=->]{muA}{A}
\ncline[arrows=->]{sigmaA}{A}

\ncline[arrows=->]{kSigma}{sigma}
\ncline[arrows=->]{lambdaSigma}{sigma}

\ncline[arrows=->]{kEpsilon}{epsilon}
\ncline[arrows=->]{lambdaEpsilon}{epsilon}

\ncline[arrows=->]{mub}{b}
\ncline[arrows=->]{sigmab}{b}

\ncline[arrows=->]{sigma}{sigmar}
\ncline[arrows=->]{epsilon}{epsilonr}

%\ncline[arrows=->]{muA}{dmuAOUT}
%\ncline[arrows=->]{kSigma}{dkSigmaOUT}
%\ncline[arrows=->]{mub}{dmubOUT}

% 4. Plates
%%%%%%%%%%%

\rput(4.25,4.25){\GM@plate{4.5}{2.5}{repeat $r=1$:$2$   }}
\rput(3.25,2.5){\GM@plate{7.5}{5.25}{game $g=1$:$G$   }}
\rput(2.25,2){\GM@plate{9.5}{7.25}{subject $s=1$:$S$   }}
\rput(1.25,1.5){\GM@plate{11.5}{9.25}{condition $i=1$:$N$   }}
%\rput(1,1.5){\GM@plate{10}{10}{$u\in\left\{ [1\mbox{ }3], [2\mbox{ }2]\right\}$   }}

% 5. Model specification
%%%%%%%%%%%%%%%%%%%%%%%%

% priors
\rput(12.75,10.5){\pnode{priors}}\GM@label{priors}{Priors}
\rput(12.75,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^A_{i} \sim \mbox{Exponential}(0.01)$}
\rput(12.75,9.5){\pnode{mub}}\GM@label{mub}{$\mu^b_{i} \sim \mbox{Gaussian}(0, 100)$, $\sigma^b_{i} \sim \mbox{Exponential}(0.01)$}
\rput(12.75,9){\pnode{kSigma}}\GM@label{kSigma}{$k^{ran}_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^{ran}_{i} \sim \mbox{Exponential}(10)$}
\rput(12.75,8.5){\pnode{kEpsilon}}\GM@label{kEpsilon}{$k^{det}_{i} \sim \mbox{Exponential}(0.01)$, $\lambda^{det}_{i} \sim \mbox{Exponential}(10)$}

% subject level params
\rput(12.75,8){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
\rput(12.75,7.5){\pnode{A}}\GM@label{A}{$A_{is} \sim \mbox{Gaussian}(\mu^A_{i}, \sigma^A_{i})$}
\rput(12.75,7){\pnode{b}}\GM@label{b}{$B_{is} \sim \mbox{Gaussian}(\mu^B_{i}, \sigma^B_{i})$}
\rput(12.75,6.5){\pnode{sigma}}\GM@label{sigma}{$\sigma^{ran}_{is} \sim \mbox{Gamma}(k^{ran}_{i}, \lambda^{ran}_{i})$}
\rput(12.75,6){\pnode{epsilon}}\GM@label{epsilon}{$\sigma^{det}_{is} \sim \mbox{Gamma}(k^{det}_{i}, \lambda^{det}_{i})$}

% game level params
\rput(12.75,5){\pnode{game}}\GM@label{game}{Deterministic noise for repeated game}
\rput(12.75,4.5){\pnode{epsilonr}}\GM@label{epsilonr}{$n^{det}_{isg} \sim \mbox{Logistic}(0, \sigma^{det}_{is})$}

% repeated game level params
\rput(12.75,4){\pnode{repeat}}\GM@label{repeat}{Random noise for each game}
\rput(12.75,3.5){\pnode{sigmar}}\GM@label{sigmar}{$n^{ran}_{isgr} \sim \mbox{Logistic}(0, \sigma^{ran}_{is})$}

% choices!
%\rput(9,5.5){\pnode{c}}
%\GM@label{c}{$p \leftarrow }  $}

\rput(12.75,2.5){\pnode{choices}}\GM@label{choices}{Observed choices}
\rput(12.75,2){\pnode{q}}
\GM@label{q}{$\Delta Q_{isgr} \leftarrow
\Delta R_{isg}+A_{is}\Delta I_{isg} + b_{is} + n^{ran}_{isgr} + n^{det}_{isg}$}
\rput(12.75,1.5){\pnode{c}}
\GM@label{c}{$c_{isgr} \sim \mbox{Bernoulli}
\left( Q_{isgr} > 0 \right)$}

% 6. End document
%%%%%%%%%%%%%%%%%

\end{pspicture}

%\endTeXtoEPS
\end{document}