### 2025-03-09 codes_original_submission/GRAPHICALMODEL/OlderFiles/horizonTask_graphicalModel_v3.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Template graphical model %
% Author: Tom Lodewyckx    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% 0. Begin document
%%%%%%%%%%%%%%%%%%%

\documentclass[12pt]{article}
\usepackage{pst-all}
\usepackage{com.braju.graphicalmodels}
%\usepackage[pass,paperwidth=8.5in,paperheight=11in]{geometry}
\usepackage{ifpdf}
\ifpdf
  \pdfpaperheight=6in
  \pdfpaperwidth=9.5in
\else
  \special{papersize=7.6in,5in}
\fi
\catcode`\@=11%
\pagestyle{empty}
%\setlength{\textheight}{9.3in}
%\setlength{\textwidth}{7in}
%\setlength{\footskip}{0.5in}
%\setlength{\topmargin}{-0.5in}
%\setlength{\headheight}{0.2in}
%\setlength{\headsep}{0in}
%\setlength{\parindent}{1pc}
%\setlength{\oddsidemargin}{-0.25in}
%\setlength{\evensidemargin}{-0.25in}
\begin{document}
\TeXtoEPS
\psset{unit=14mm,arrowscale=1.5}
\SpecialCoor

% 1. Grid
%%%%%%%%%

\begin{pspicture}(2.95,2)(13,9)
%\showgrid

% 2. Nodes
%%%%%%%%%%

%\rput(1,1){\GM@node[NodeOptions]{name}}
%                 \GM@label[LabelOptions]{name}{label}

% dm and dI
\rput(3,4){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{dm}}\GM@label[offset=-5.7mm]{dm}{$\Delta R_{nst}$}
\rput(5,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dI}}\GM@label[offset=-5mm]{dI}{$\Delta I_{nst}$}
\rput(7,4){\GM@node[query=true,observed=true,observedColor=lightgray,nodeSize=12mm]{dl}}\GM@label[offset=-5.7mm]{dl}{$\Delta l_{nst}$}

% choices
\rput(5,6){\GM@node[observed=true,observedColor=lightgray,nodeSize=12mm]{c}}\GM@label[offset=-3mm]{c}{$c_{nst}$}

% subject parameters
\rput(3,8){\GM@node[nodeSize=12mm]{A}}\GM@label[offset=-3mm]{A}{$A_{ns}$}
\rput(5,8){\GM@node[nodeSize=12mm]{b}}\GM@label[offset=-3mm]{b}{$B_{ns}$}
\rput(7,8){\GM@node[nodeSize=12mm]{sigma}}\GM@label[offset=-2mm]{sigma}{$\sigma_{ns}$}

% hyperparameters
\rput(2,10){\GM@node[nodeSize=12mm]{muA}}\GM@label[offset=-2mm]{muA}{$\mu^A_{n}$}
\rput(3,10){\GM@node[nodeSize=12mm]{sigmaA}}\GM@label[offset=-2mm]{sigmaA}{$\sigma^A_{n}$}

\rput(4.5,10){\GM@node[nodeSize=12mm]{mub}}\GM@label[offset=-2mm]{mub}{$\mu^B_{n}$}
\rput(5.5,10){\GM@node[nodeSize=12mm]{sigmab}}\GM@label[offset=-2mm]{sigmab}{$\sigma^B_{n}$}

\rput(7,10){\GM@node[nodeSize=12mm]{kSigma}}\GM@label[offset=-2mm]{kSigma}{$k^\sigma_{n}$}
\rput(8,10){\GM@node[nodeSize=12mm]{lambdaSigma}}\GM@label[offset=-2mm]{lambdaSigma}{$\lambda^\sigma_{n}$}

% differences
%\rput(3,13){\GM@node[nodeSize=12mm]{dmuAOUT}}
%\rput(3,13){\GM@node[nodeSize=10mm]{dmuA}}\GM@label[offset=-3mm]{dmuA}{$\delta\mu^A_{u}$}
%\rput(5,13){\GM@node[nodeSize=12mm]{dkSigmaOUT}}
%\rput(5,13){\GM@node[nodeSize=10mm]{dkSigma}}\GM@label[offset=-3mm]{dkSigma}{$\delta k^\sigma_{u}$}
%\rput(7,13){\GM@node[nodeSize=12mm]{dmubOUT}}
%\rput(7,13){\GM@node[nodeSize=10mm]{dmub}}\GM@label[offset=-3mm]{dmub}{$\delta\mu^b_{u}$}

% 3. Arrows
%%%%%%%%%%%
%
\ncline[arrows=->]{dm}{c}
\ncline[arrows=->]{dI}{c}
\ncline[arrows=->]{dl}{c}

\ncline[arrows=->]{A}{c}
\ncline[arrows=->]{sigma}{c}
\ncline[arrows=->]{b}{c}

\ncline[arrows=->]{muA}{A}
\ncline[arrows=->]{sigmaA}{A}

\ncline[arrows=->]{kSigma}{sigma}
\ncline[arrows=->]{lambdaSigma}{sigma}

\ncline[arrows=->]{mub}{b}
\ncline[arrows=->]{sigmab}{b}

%\ncline[arrows=->]{muA}{dmuAOUT}
%\ncline[arrows=->]{kSigma}{dkSigmaOUT}
%\ncline[arrows=->]{mub}{dmubOUT}

% 4. Plates
%%%%%%%%%%%

\rput(2.25,3){\GM@plate{5.5}{4}{game $g=1$:$G$   }}
\rput(1.75,2.5){\GM@plate{6.5}{6.5}{subject $s=1$:$S$   }}
\rput(1.25,2){\GM@plate{7.5}{8.75}{condition $n=1$:$N$   }}
%\rput(1,1.5){\GM@plate{10}{10}{$u\in\left\{ [1\mbox{ }3], [2\mbox{ }2]\right\}$   }}

% 5. Model specification
%%%%%%%%%%%%%%%%%%%%%%%%

% priors
\rput(9,10.5){\pnode{priors}}\GM@label{priors}{Priors}
\rput(9,10){\pnode{muA}}\GM@label{muA}{$\mu^A_{n} \sim \mbox{Gaussian}(0, 1000)$}
\rput(9,9.5){\pnode{sigmaA}}\GM@label{sigmaA}{$\sigma^A_{n} \sim \mbox{Gamma}(0.001, 0.001)$}

\rput(9,9){\pnode{mub}}\GM@label{mub}{$\mu^B_{n} \sim \mbox{Gaussian}(0, 1000)$}
\rput(9,8.5){\pnode{sigmab}}\GM@label{sigmab}{$\sigma^B_{n} \sim \mbox{Gamma}(0.001, 0.001)$}

\rput(9,8){\pnode{kSigma}}\GM@label{kSigma}{$k^\sigma_{n} \sim \mbox{Exponential}(0.001)$}
\rput(9,7.5){\pnode{lambdaSigma}}\GM@label{lambdaSigma}{$\lambda^\sigma_{n} \sim \mbox{Exponential}(0.001)$}

% subject level params
\rput(9,6.5){\pnode{subpars}}\GM@label{subpars}{Subject specific parameters}
\rput(9,6){\pnode{A}}\GM@label{A}{$A_{ns} \sim \mbox{Gaussian}(\mu^A_{n}, \sigma^A_{n})$}
\rput(9,5.5){\pnode{b}}\GM@label{b}{$B_{ns} \sim \mbox{Gaussian}(\mu^B_{n}, \sigma^B_{n})$}
\rput(9,5){\pnode{sigma}}\GM@label{sigma}{$\sigma_{ns} \sim \mbox{Gamma}(k^\sigma_{n}, \lambda^\sigma_{n})$}

% choices!
%\rput(9,5.5){\pnode{c}}
%\GM@label{c}{$p \leftarrow }  $}

\rput(9,4){\pnode{choices}}\GM@label{choices}{Observed choices}
\rput(9,3.5){\pnode{p}}
\GM@label{p}{$p_{nst} \leftarrow
\left[1+\exp\left( \frac{\Delta R_{nst}+A_{ns}\Delta I_{nst} + B_{ns}\Delta l_{nst}}{\sigma_{ns}} \right)\right]^{-1} $}
\rput(9,2.9){\pnode{c}}
\GM@label{c}{$c_{nst} \sim \mbox{Bernoulli}
\left( p_{nst} \right)$}

% 6. End document
%%%%%%%%%%%%%%%%%

\end{pspicture}

\endTeXtoEPS
\end{document}